\documentclass[11pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{amssymb}

% Page style
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.4pt}
\fancyhead[L]{CPF Authorized Service Provider Agreement}
\fancyhead[R]{Provider \#: \_\_\_\_\_\_\_\_}
\fancyfoot[C]{\thepage}

% Spacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\title{\textbf{CPF AUTHORIZED SERVICE PROVIDER\\CERTIFICATION AGREEMENT}}
\author{}
\date{}

\begin{document}

\maketitle

\section*{PARTIES}

This Authorized Service Provider Certification Agreement ("Agreement") is entered into as of the \_\_\_ day of \_\_\_\_\_\_\_\_\_\_\_, 20\_\_\_ ("Effective Date"), by and between:

\textbf{[CERTIFICATION BODY NAME]} ("Certification Body" or "CB")\\
Authorized CPF Certification Body\\
Address: [Address]\\
Email: [Email]

AND

\textbf{[FIRM NAME]} ("Firm" or "Authorized Service Provider")\\
A [jurisdiction] [entity type]\\
Registration: [Number]\\
Address: [Address]\\
Website: [Website]\\
Email: [Email]

\section*{RECITALS}

WHEREAS, CB is authorized by CPF3 to certify organizations as Authorized Service Providers;

WHEREAS, Firm provides CPF-related services and desires "CPF Authorized Service Provider" certification;

WHEREAS, Firm employs certified CPF professionals and maintains quality management systems;

NOW, THEREFORE, the Parties agree as follows:

\section{DEFINITIONS}

\textbf{1.1 "Authorized Service Provider"} means a firm certified to provide CPF services and use the "CPF Authorized Service Provider" designation.

\textbf{1.2 "CPF Services"} means professional services including assessments, implementation, training, monitoring, and integration.

\textbf{1.3 "Certified Personnel"} means individuals with valid CPF certifications employed by Firm.

\textbf{1.4 "ASP Mark"} means the "CPF Authorized Service Provider" trademark and logo.

\section{FIRM SIZE CLASSIFICATION}

\begin{itemize}
\item[$\square$] Micro (1-10 employees)
\item[$\square$] Small (11-50 employees)
\item[$\square$] Medium (51-250 employees)
\item[$\square$] Large (250+ employees)
\end{itemize}

Total: \underline{\hspace{4cm}}

\section{REQUIREMENTS}

\textbf{3.1 Staffing:}

\begin{tabular}{|l|c|p{5cm}|}
\hline
Size & Min. Certified & Composition \\
\hline
Micro & 2 & Any combination \\
Small & 5 & Min. 2 Assessors \\
Medium & 10 & Min. 5 Assessors + 1 Auditor \\
Large & 20 & Min. 10 Assessors + 2 Auditors \\
\hline
\end{tabular}

All must:
\begin{itemize}
\item Hold current CPF certifications
\item Be employed or exclusive contract (1+ year)
\item Active in CPF services (25\%+ time)
\item Maintain CPE requirements
\end{itemize}

\textbf{3.2 Quality Management:}

\textit{SOPs:}
\begin{itemize}
\item Assessment methodology
\item Scoring guidelines
\item Privacy protocols
\item Report standards
\item Client communication
\item Project management
\end{itemize}

\textit{Quality Review:}
\begin{itemize}
\item Peer review all reports
\item Quality checklists
\item Client satisfaction surveys
\item Quarterly metrics review
\item Root cause analysis
\item CAPA system
\end{itemize}

\textit{Metrics:}
\begin{itemize}
\item Client satisfaction >4.0/5.0
\item Delivery timeliness
\item Complaint rate <5\%
\item Staff utilization
\item CPE completion
\end{itemize}

\textbf{3.3 Privacy and Ethics:}

\textit{Privacy:}
\begin{itemize}
\item Privacy policy
\item Differential privacy ($\varepsilon \leq 0.1$)
\item Minimum aggregation (10)
\item Encryption (AES-256, TLS 1.3)
\item MFA and access controls
\item Audit logging
\item Retention policy (max 5 years)
\item Annual external privacy audit
\end{itemize}

\textit{Ethics:}
\begin{itemize}
\item CPF Code of Ethics adoption
\item Annual training (2 hours)
\item Complaint procedures
\item Conflict management
\item Independence protocols
\end{itemize}

\textbf{3.4 Insurance:}

\textit{Professional Liability:}

\begin{tabular}{|l|c|c|}
\hline
Size & Per Occurrence & Aggregate \\
\hline
Micro/Small & \$1M & \$2M \\
Medium & \$2M & \$4M \\
Large & \$5M & \$10M \\
\hline
\end{tabular}

\textit{Cyber:} \$1M all sizes

CB and CPF3 additional insureds.

\section{CERTIFICATION PROCESS}

\textbf{4.1 Application (Weeks 1-4):}

Submit:
\begin{itemize}
\item Application form
\item Certified personnel list
\item QMS documentation
\item Privacy framework and audit
\item Ethics policies and records
\item Insurance certificates
\item Sample deliverables (3 redacted)
\item Client references (5)
\item Org chart
\item Application fee
\end{itemize}

CB conducts:
\begin{itemize}
\item Desktop review
\item Personnel verification
\item Reference checks (3 minimum)
\item QMS assessment
\end{itemize}

\textbf{4.2 Audit (Weeks 6-8):}

Duration: 2-4 days

Activities:
\begin{itemize}
\item Opening meeting
\item QMS review
\item Personnel interviews
\item Project file review (3-5)
\item Infrastructure assessment
\item Privacy evaluation
\item Ethics assessment
\item Insurance verification
\item Closing meeting
\end{itemize}

Findings: Conformity, Minor NC, Major NC, Observation

Report: 15 business days

\textbf{4.3 Corrective Actions:}

If NCs:
\begin{itemize}
\item Plan within 30 days
\item Major: Before certification
\item Minor: Within 90 days after
\item Verification
\end{itemize}

\textbf{4.4 Decision (15 days):}

\textit{Granted:}
\begin{itemize}
\item Certificate (electronic 3 days, physical 10)
\item Digital badge
\item Featured directory listing
\item Marketing toolkit
\item Resources access
\item Surveillance schedule
\end{itemize}

\textit{Denied:}
\begin{itemize}
\item Written explanation
\item Remediation guidance
\item Appeal rights (30 days)
\item Reapplication option
\end{itemize}

\section{BENEFITS}

\textbf{5.1 Marketing:}
\begin{itemize}
\item Use ASP trademark
\item Official logo (multiple formats)
\item Digital badge
\item Marketing materials authorization
\item Approved messaging
\end{itemize}

\textbf{5.2 Directory:}
\begin{itemize}
\item Priority placement
\item Enhanced profile
\item Logo, description, services
\item Team size indicators
\item Analytics
\end{itemize}

\textbf{5.3 Business Development:}
\begin{itemize}
\item CPF3 referrals
\item Preferred provider status
\item RFP access
\item Reduced due diligence
\item Premium pricing justification
\end{itemize}

\textbf{5.4 Resources:}
\begin{itemize}
\item Exclusive tools and templates
\item Early access (90 days)
\item Dedicated account manager
\item Priority support
\item Volume discounts: 10-20\%
\item Free webinars
\item Conference passes
\end{itemize}

\textbf{5.5 Community:}
\begin{itemize}
\item Annual Provider Summit
\item Quarterly roundtables
\item Private forum
\item Regional chapters
\end{itemize}

\section{ONGOING OBLIGATIONS}

\textbf{6.1 Annual Update (30 days before anniversary):}
\begin{itemize}
\item Personnel roster
\item Quality metrics
\item Projects completed
\item Client satisfaction
\item Insurance certificates
\item Privacy audit results
\item Ethics training records
\item Material changes
\end{itemize}

\textbf{6.2 Bi-annual Surveillance (1-3 days):}
\begin{itemize}
\item Focused review
\item Sample projects (2-3)
\item Personnel subset
\item Continued conformity
\item Changes verification
\end{itemize}

\textbf{6.3 Performance Monitoring:}

CB monitors:
\begin{itemize}
\item Minimum personnel maintenance
\item Client satisfaction (>4.0/5.0)
\item Complaint rates (<5\%)
\item Insurance status
\item Privacy audits
\end{itemize}

\textbf{6.4 Change Notifications:}

\textit{Immediate (5 days):}
\begin{itemize}
\item Below minimum personnel
\item Insurance lapse
\item Privacy/data breach
\item Ethics complaints
\item Legal actions
\item Major personnel loss
\end{itemize}

\textit{30-day:}
\begin{itemize}
\item Ownership changes
\item Reorganizations
\item Name changes
\item Office changes
\end{itemize}

\section{SUSPENSION AND REVOCATION}

\textbf{7.1 Suspension (max 180 days):}

Grounds:
\begin{itemize}
\item Below minimum personnel
\item Insurance lapse
\item Surveillance failure
\item Fee non-payment
\item Complaints under investigation
\item Privacy breach
\item Major NC not corrected (90 days)
\item Client satisfaction <3.5 (2 quarters)
\end{itemize}

Effect:
\begin{itemize}
\item Restriction on new Mark use
\item Directory: "Suspended"
\item Add "Status Under Review"
\item Referrals suspended
\item 180-day remediation or revocation
\end{itemize}

\textbf{7.2 Revocation:}

Grounds:
\begin{itemize}
\item Failure to remediate
\item Severe privacy violations
\item Fraud/misrepresentation
\item Systematic quality failures
\item Ethics violations
\item Persistent Mark misuse
\item Insurance loss >60 days
\item Dissolution/bankruptcy
\end{itemize}

Process:
\begin{itemize}
\item Written notice
\item 30 days to respond
\item Independent review
\item Decision within 45 days
\item Appeal rights
\end{itemize}

Effect:
\begin{itemize}
\item Immediate cessation ALL Mark use
\item Directory removal
\item Public notice (12 months)
\item No refunds
\item Reapplication: 2-5 years or permanent
\item Must notify clients
\end{itemize}

\textbf{7.3 Voluntary Withdrawal:}
\begin{itemize}
\item 60 days notice
\item No negative record
\item Reapply after 12 months
\item Simplified if within 24 months
\end{itemize}

\section{FEES}

\textbf{8.1 Application:}

\begin{tabular}{|l|c|}
\hline
Size & Fee \\
\hline
Micro & \$1,000 \\
Small & \$2,000 \\
Medium & \$3,500 \\
Large & \$5,000 \\
\hline
\end{tabular}

\textbf{8.2 Audit:}

\begin{tabular}{|l|c|}
\hline
Size & Fee \\
\hline
Micro & \$3,000 \\
Small & \$6,000 \\
Medium & \$10,000 \\
Large & \$15,000 \\
\hline
\end{tabular}

Additional: \$1,500/day

\textbf{8.3 Certification:}

\begin{tabular}{|l|c|}
\hline
Size & Fee \\
\hline
Micro & \$1,500 \\
Small & \$2,500 \\
Medium & \$4,000 \\
Large & \$6,000 \\
\hline
\end{tabular}

\textbf{8.4 Annual Surveillance:} Same as certification

\textbf{8.5 Recertification (3 years):}
\begin{itemize}
\item Audit: 75\% of initial
\item Fee: Same as initial
\end{itemize}

\textbf{8.6 Other:}
\begin{itemize}
\item Follow-up: \$1,500/day
\item Expedited: 25\% surcharge
\item Travel: Actual costs
\end{itemize}

\textbf{8.7 Payment:}
\begin{itemize}
\item Application: With submission
\item Audit: Before scheduling
\item Certification: Upon decision
\item Surveillance: 30 days before anniversary
\item USD, 30-day terms
\item Late: 1.5\% monthly
\item Suspended if >60 days overdue
\end{itemize}

\section{APPEALS}

\textbf{9.1 Appealable:}
\begin{itemize}
\item Denial
\item Suspension
\item Revocation
\item Major NC findings
\end{itemize}

\textbf{9.2 Process:}
\begin{itemize}
\item Written within 30 days
\item Fee: \$500
\item Grounds and evidence
\item Independent panel
\item Decision within 45 days
\item Uphold/Modify/Reverse/Remand
\item Fee refunded if successful
\item Final and binding
\end{itemize}

\section{CONFIDENTIALITY}

\textbf{10.1 CB Maintains:}
\begin{itemize}
\item QMS documentation
\item Client lists
\item Business information
\item Audit findings
\item Not disclosed except: Public info, to CPF3, to accreditation bodies, as required by law
\end{itemize}

\textbf{10.2 Firm Maintains:}
\begin{itemize}
\item All client information
\item Assessment data
\item Client identities
\item CB audit methodologies
\end{itemize}

\textbf{10.3 Data Protection:}
\begin{itemize}
\item GDPR/CCPA compliance
\item Security measures
\item Breach notification (24 hours)
\item Cooperation
\end{itemize}

\section{LIMITATION OF LIABILITY}

\textbf{11.1 Disclaimer.} NO WARRANTIES REGARDING BUSINESS OUTCOMES, CLIENT ACQUISITION, REVENUE, OR COMPETITIVE ADVANTAGES.

\textbf{11.2 Limitation.} NO LIABILITY FOR INDIRECT, CONSEQUENTIAL, SPECIAL, OR PUNITIVE DAMAGES.

\textbf{11.3 Cap.} NOT TO EXCEED FEES PAID IN 12 MONTHS PRECEDING CLAIM.

\textbf{11.4 Exceptions:} Gross negligence, confidentiality breaches, data violations, indemnification, claims not limitable by law.

\section{INDEMNIFICATION}

\textbf{12.1 By Firm:} From claims arising from services to clients, negligence, Mark misuse, misrepresentation, privacy violations, false information, third-party claims.

\textbf{12.2 By CB:} From confidentiality breach, audit negligence, data violations.

\section{GENERAL PROVISIONS}

\textbf{13.1 Governing Law.} [Jurisdiction]

\textbf{13.2 Disputes.} Negotiation (30 days), mediation, then arbitration.

\textbf{13.3 Entire Agreement.} This Agreement and Schedules.

\textbf{13.4 Amendment.} CB may amend (180 days notice).

\textbf{13.5 Assignment.} Firm cannot without consent; CB may for business transfer.

\textbf{13.6 Force Majeure.} Neither liable for events beyond control (>90 days).

\textbf{13.7 Notices.} Written to stated addresses.

\textbf{13.8 Severability.} Invalid provisions reformed.

\textbf{13.9 Independent Contractor.} Parties independent, not partners.

\textbf{13.10 Survival.} Sections 6.3, 10, 11, 12, 13 survive.

\textbf{13.11 CPF3 Relationship:}
\begin{itemize}
\item CPF3 owns CPF IP
\item CB operates under CPF3 license
\item Firm has no direct CPF3 relationship
\item CPF3 third-party beneficiary
\end{itemize}

\section{ACKNOWLEDGMENTS}

Firm acknowledges:
\begin{itemize}
\item Read and understands Agreement
\item Agrees to CPF Code of Ethics
\item Provided accurate information
\item Understands requirements and obligations
\item Fees non-refundable
\item May be suspended/revoked
\item Must maintain all requirements
\item Authorizes verification
\item Authorizes directory publication
\item Consents to data processing
\item No guarantee of business outcomes
\item Will cease Mark use if certification ends
\item Agrees to arbitration
\end{itemize}

\section*{SIGNATURES}

\textbf{CERTIFICATION BODY:}

By: \underline{\hspace{6cm}} Date: \underline{\hspace{3cm}}

Name: \underline{\hspace{6cm}} Title: \underline{\hspace{6cm}}

\vspace{2em}

\textbf{FIRM:}

By: \underline{\hspace{6cm}} Date: \underline{\hspace{3cm}}

Name: \underline{\hspace{6cm}} Title: \underline{\hspace{6cm}}

\vspace{2em}

\section*{CERTIFICATION RECORD}

\begin{tabular}{|l|p{10cm}|}
\hline
Firm Size & \\
\hline
Certificate Number & \\
\hline
Issue Date & \\
\hline
Expiration Date & \\
\hline
Certified Personnel at Certification & \\
\hline
Issued By & \\
\hline
\end{tabular}

\vspace{2em}

\begin{center}
\textit{End of Authorized Service Provider Certification Agreement}
\end{center}

\end{document}